### 2025-05-27 Manuscript/Lower bounds of the success probability in quantum state exclusion for general ensembles.tex
Writthing

\documentclass[12pt,letterpaper]{article}
\usepackage{fullpage}
\usepackage[top=1.5cm, bottom=3.5cm, left=2.2cm, right=2.2cm]{geometry}
\usepackage{amsmath,amsthm,amsfonts,amssymb,amscd, esint}
\usepackage{lastpage}
\usepackage{enumerate}
\usepackage{fancyhdr}
\usepackage{mathrsfs}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{hyperref}
\usepackage[english]{babel}
\usepackage{lipsum}
\usepackage[table,xcdraw]{xcolor}
\usepackage{enumitem}
\usepackage{overpic}
\usepackage{float}
\usepackage{chemfig}
\usepackage{yfonts}
\usepackage{braket}
\usepackage{dsfont}
\usepackage{tikz}
\usepackage{wrapfig}
\usepackage{url}
\usepackage{natbib}
\usepackage[normalem]{ulem}
\usepackage{multicol}
\useunder{\uline}{\ul}{}

@@

% --- Section formatting ---
\titleformat{\section}
  {\normalfont\normalsize\bfseries\centering} % font size, bold, centered
  {\thesection}{1em}{\MakeTextUppercase} % uppercase text

@@

% --- Roman numerals for numbering ---
\renewcommand{\thesection}{\Roman{section}}
\renewcommand{\thesubsection}{\Roman{section}.\Roman{subsection}}

@@

%\setlength{\parindent}{0.0in}
%\setlength{\parskip}{0.05in}

% Edit these as appropriate
\newcommand\course{}
\newcommand\subject{Final Degree Project}
\newcommand\degree{Bachelor's Degree in Physics}
\newcommand\documenttitle{Lower bounds of the success probability in quantum state exclusion for general ensembles}
\newcommand\NetIDb{Universitat Autònoma de Barcelona}
\newcommand\AuthorName{Sergio Castañeiras Morales}

@@

\begin{document}
\title{\vspace{4cm} \normalsize 
		\includegraphics[width = 0.25\textwidth]{GeneralSources/UABLogo.png}\\ [0.5cm]
		\textsc{\NetIDb}\\ [2.0cm]
		\HRule{0.5pt} \\
		\LARGE \textbf{\uppercase{\documenttitle}}
		\HRule{2pt} \\ [1.5cm]
		\normalsize \begin{tabular}{rcl}  % Create a right-left column alignment
        \textsc{Author} & : & \textsc{\AuthorName} \\
        \textsc{Supervisor} & : & \textsc{Ramón Muñoz Tapia} \\
        \textsc{Co-Supervisor} & : & \textsc{Santiago Llorens Fernández}
    \end{tabular}
    \normalsize \vspace*{5\baselineskip}
		}

@@

\thispagestyle{empty}
\mbox{} 
\newpage
\thispagestyle{empty}
\vspace*{\fill} % Push content to vertical center
\begin{flushright}
    \emph{Ab ovo usque ad mala.}\\[1em]
    \textbf{Horace}
\end{flushright}
\vspace*{\fill} 
\newpage
\thispagestyle{empty}
\mbox{} 
\newpage

\newpage
\pagestyle{fancyplain}
\headheight 35pt
\lhead{\NetIDb}    
\rhead{\subject}
\cfoot{\small\thepage}
\headsep 1.5em
\setcounter{page}{1}
\begin{multicols}{2}[\begin{center}
\begin{abstract}
Given a quantum state known to be prepared from an ensemble of two or more states, quantum state exclusion aims to rule out the possibility that it was prepared in a particular state from the ensemble. Using the known solution for group generated ensembles \cite{MainPaper}, we study this result as a lower bound for randomly generated ensembles via semidefinite programming.
\end{abstract}
\end{center}
Keywords: \emph{SDP, Quantum state exclusion , \textcolor{blue}{Add keywords}}.
]
\section{Introduction}
In many real-world scenarios, excluding a certain hypothesis can be more practical than solving the problem entirely. For instance, in disease diagnosis, ruling out potential diseases often serves as the first step in identifying the actual condition. Similarly, when repairing a machine, it is sometimes more efficient to identify the components that are functioning correctly, which narrows down the search for the faulty part.

In this project, we project this idea into the quantum realm by focusing on \gls{qse}. Rather than determining the exact state of a quantum system, we aim to eliminate one or more possible candidates from a known ensemble of states. Notice, this approach can be more suitable or efficient in certain quantum information tasks.

Given a quantum state known to be prepared from a finite ensemble, \gls{qsd} seeks to identify which specific state from the ensemble corresponds to the given system. In contrast, \gls{qse}~\cite{PhysicsExclusionSource} adopts the opposite perspective, it aims to determine which states from the ensemble do not correspond to the prepared state. While \gls{qsd} has been deeply studied in recent years\cite{DiscriminationArticle}, with significant advances since its inception~\cite{helstromBook}, \gls{qse} offers a complementary framework with distinct advantages.

Although the tasks of exclusion and discrimination coincide for ensembles containing only two states \footnote{Since for the two states case excluding one necessarily implies identifying the other.}, when dealing with ensembles of three or more states, the two problems diverge in both approach and complexity. One of the most significant features of \gls{qse} is the possibility of achieving \emph{perfect exclusion}, where certain states can be ruled out with zero probability of error in cases where \emph{perfect discrimination} is impossible\cite{OptimalitySRM}.

This capability opens new frontiers in quantum information theory, particularly in the context of partial information retrieval from quantum systems. By excluding certain states, it is possible to gain insight into the encoded information without needing to fully determine the original state.

As with \gls{qsd}, obtaining a general analytical solution for \gls{qse} remains an open problem. However, analytical results have been found in specific cases when the ensemble of quantum states exhibits a certain degree of symmetry. In particular, when the ensemble is generated by the action of a finite group, the problem becomes more tractable and exact solutions have been derived.

The exclusion task can be carried out under two main protocols, \gls{me} and \gls{ze}\footnote{Also known as \emph{unambiguous exclusion}.}. In the Minimum Error scenario, the goal is to minimize the probability of mistakenly excluding the actual prepared state. In contrast, the Zero Error approach seeks to exclude a state with absolute certainty, even if that means sometimes the procedure yields an inconclusive result.

Building on recent results that provide exact solutions for exclusion tasks in group generated ensembles \cite{MainPaper}, this project undertakes a numerical study of such results as lower bounds for more general, randomly generated ensembles. To this end, we employ \gls{sdp} to explore \gls{qse} performance in arbitrary settings. Furthermore, we investigate improved bounds for the general case based on how closely a given ensemble resembles a group generated one\footnote{The notion of "how close" will be formally defined in Section~\textcolor{blue}{add section}.}.

\subsection{Preamble}

This project is entirely situated within the quantum realm. Therefore, we begin by contextualizing some foundational postulates, notation, and elements upon which the rest of the project relies.

@@

Note that if $\ket{\psi'} = e^{i\theta}\ket{\psi}$ for some global phase $\theta \in \mathbb{R}$, the observable outcomes and probabilities remain unchanged. Therefore, quantum states differing only by a global phase are physically indistinguishable.

In relation to Postulate I, it is often convenient to represent quantum states using the \emph{density matrix} formalism, where the state is described by $\rho = \ket{\psi}\bra{\psi}$ instead of $\ket{\psi}$. In this formalism, the global phase ambiguity disappears, since if $\ket{\psi'} = e^{i\theta}\ket{\psi}$, then
\begin{align*}
	\rho' = \ket{\psi'}\bra{\psi'} = \ket{\psi}\bra{\psi} = \rho.
\end{align*}

@@

Both formulations are equivalent, though one may be more convenient than the other depending on the context. Furthermore, the density matrix formalism naturally introduces the concept of \emph{mixed states}. Until now, we have implicitly dealt only with \emph{pure states}, i.e., states of the form $\rho = \ket{\psi}\bra{\psi}$.

Suppose now that, for reasons we ignore, the system is prepared in one of several pure states $\{\rho_i\}_{i=1}^n$ with probabilities $\{\eta_i\}_{i=1}^n$. Then the overall state of the system is described by the mixed state
\begin{align*}
	\rho = \sum_{i=1}^n \eta_i \rho_i.
\end{align*}

@@

In relation to the observable postulate, we now define a \gls{povm} as a collection of measurement operators $\{O_i\}_{i=1}^n$ that satisfy,
\begin{itemize}
	\item Each $O_i$ is positive semi-definite, meaning all its eigenvalues $o_j^{(i)}$ satisfy $o_j^{(i)} \geq 0$.
	\item The operators satisfy the completeness relation,
\begin{align*}
	\sum_{i=1}^n O_i = \mathds{1},
\end{align*}
where $\mathds{1}$ denotes the identity operator on the Hilbert space $\mathcal{H}$.
\end{itemize}

Unlike projective measurements, the operators $O_i$ in a POVM are not necessarily orthogonal projectors. This allows the number of measurement outcomes to exceed the dimensionality of the Hilbert space, making POVMs particularly useful in tasks such as \gls{qse} and \gls{qsd}.

\subsection{Formulation of the Problem}\label{sectionFormulationOfTheProblem}

Let $\left\{(\rho_i, \eta_i)\right\}_{i=1}^n$ be an ensemble of $n$ quantum states, where each $\rho_i$ denotes a pure state density matrix\footnote{This formulation hols true for mixed states but the project will only discuss the pure state scenario.}, i.e., $\rho_i = \ket{\psi_i}\bra{\psi_i}$, and $\eta_i$ represents the prior probability of occurrence of the state $\rho_i$. Let $\rho_j$ be the target state from this ensemble. Our objective is to develop a procedure to identify another state $\rho_k \in \left\{\rho_i\right\}_{i=1}^n$, such that $\rho_k \neq \rho_j$.

Quantum measurements are described by a set of \glspl{povm}, denoted by $\left\{\Pi_i\right\}_{i=1}^n$, acting on the Hilbert space $\mathcal{H}$ of the quantum system. Here we pewsent the two studied protocols for \gls{qse}: minimum-error (\gls{me}) and zero-error (\gls{ze}).

The goal of the \gls{me} protocol is to minimize the probability of incorrectly excluding the target state from our hypothesis. If we formulate it as an \gls{sdp}, the problem reads,\footnote{Note that the \gls{sdp} formulations of quantum state discrimination may differ from the exclusion ones by interchanging minimization and maximization problems.}
\begin{align*}
	P_{\text{\gls{me}}}^e = \min_{\left\{\Pi_i\right\}} &\sum_{i=1}^n \tr(\Pi_i \rho_i),\\
	\text{subject to} \quad & \sum_{i=1}^n \Pi_i = \mathds{1}, \quad \Pi_i \geq 0 \quad \forall i \in \{1, \dots, n\}.
\end{align*}

Note the constraints $\sum_{i=1}^n \Pi_i = \mathds{1}$ and $\Pi_i \geq 0$ ensure that the $\Pi_i$ form a valid POVM, since they demand positive semi-definition and form a complete measurement. The superscript $e$ in $P_{\text{\gls{me}}}^e$ indicates that this is the \emph{error probability}.

Alternatively, we may formulate the problem in terms of the \emph{success probability}, denoted by $P_{\text{\gls{me}}}^s$, which quantifies the probability of a correct exclusion. This equivalent formulation reads,
\begin{align*}
	P_{\text{\gls{me}}}^s = \max_{\left\{\Pi_i\right\}} &\left(1 - \sum_{i=1}^n \tr(\Pi_i \rho_i)\right),\\
	\text{subject to} \quad & \sum_{i=1}^n \Pi_i = \mathds{1}, \quad \Pi_i \geq 0 \quad \forall i \in \{1, \dots, n\}.
\end{align*}

Naturally, both formulations are related via:
\begin{align*}
P_{\text{\gls{me}}}^s + P_{\text{\gls{me}}}^e = 1.
\end{align*}

In the case of the \gls{ze} protocol, the POVMs must also satisfy an unambiguity condition, i.e. each measurement operator $\Pi_i$ must be orthogonal to the corresponding state $\rho_i$. In other words,
\begin{align*}
\tr(\Pi_i \rho_i) = 0 \quad \forall i \in \{1, \dots, n\}.
\end{align*}

To ensure completeness, we introduce an additional POVM element $\Pi_?$ representing an inconclusive result,
\begin{align*}
\Pi_? = \mathds{1} - \sum_{i=1}^n \Pi_i.
\end{align*}
If the measurement yields the outcome $\Pi_?$ (i.e., the "?" symbol), the result is inconclusive.

The corresponding \gls{sdp} for minimizing the probability of an inconclusive result (i.e., error) in the \gls{ze} protocol is:
\begin{align*}
	P_{\text{\gls{ze}}}^e = \min_{\left\{\Pi_i\right\}} &\sum_{i=1}^n \tr(\Pi_? \rho_i),\\
	\text{subject to} \quad & \sum_{i=1}^n \Pi_i + \Pi_? = \mathds{1}, \quad \Pi_? \geq 0,\\
	& \tr(\Pi_i \rho_i) = 0, \quad \Pi_i \geq 0 \quad \forall i.
\end{align*}

The corresponding success probability is naturally given by,
\begin{align*}
	P_{\text{\gls{ze}}}^s = \max_{\left\{\Pi_i\right\}} &\left(1 - \sum_{i=1}^n \tr(\Pi_? \rho_i)\right),\\
	\text{subject to} \quad & \sum_{i=1}^n \Pi_i + \Pi_? = \mathds{1}, \quad \Pi_? \geq 0,\\
	& \tr(\Pi_i \rho_i) = 0, \quad \Pi_i \geq 0 \quad \forall i.
\end{align*}

This formulation is analogous to the \gls{me} protocol, with the crucial difference being the constraint $\tr(\Pi_i \rho_i) = 0$, enforcing unambiguous discrimination.

\subsection{Gram Matrix Formulation}\label{sectionGramMatrixFormulation}

Let $\mathcal{G} \in \mathbb{C}^{n \times n}$ be the \emph{Gram matrix} of the system, defined as the $n \times n$ positive semidefinite Hermitian matrix given by,
\begin{equation*}
	\mathcal{G} =
	\begin{pmatrix}
		\braket{\psi_1|\psi_1} & \braket{\psi_1|\psi_2} & \dots & \braket{\psi_1|\psi_n} \\
		\braket{\psi_2|\psi_1} & \braket{\psi_2|\psi_2} & \dots & \braket{\psi_2|\psi_n} \\
		\vdots & \vdots & \ddots & \vdots \\
		\braket{\psi_n|\psi_1} & \braket{\psi_n|\psi_2} & \dots & \braket{\psi_n|\psi_n}
	\end{pmatrix},
\end{equation*}
that is, $\mathcal{G}_{i,j} = \braket{\psi_i|\psi_j}$. Since all states are normalized, the diagonal elements are equal to 1, and the Gram matrix becomes,
\begin{equation*}
	\mathcal{G} =
	\begin{pmatrix}
		1 & \braket{\psi_1|\psi_2} & \dots & \braket{\psi_1|\psi_n} \\
		\braket{\psi_2|\psi_1} & 1 & \dots & \braket{\psi_2|\psi_n} \\
		\vdots & \vdots & \ddots & \vdots \\
		\braket{\psi_n|\psi_1} & \braket{\psi_n|\psi_2} & \dots & 1
	\end{pmatrix}.
\end{equation*}

By construction, the Gram matrix is Hermitian,
\begin{align*}
	\mathcal{G}_{i,j}^* = (\braket{\psi_i|\psi_j})^* = \braket{\psi_j|\psi_i} = \mathcal{G}_{j,i}.
\end{align*}
Additionally, $\mathcal{G}$ is positive semidefinite. To see this, consider an arbitrary state $\ket{\Phi}$ and an orthonormal basis $\{\ket{i}\}_{i=1}^n$, then,
\begin{align*}
	\braket{\Phi|\mathcal{G}|\Phi} &= \braket{\Phi| \left( \sum_{i,j=1}^{n} \braket{\psi_i|\psi_j} \ket{i}\bra{j} \right) |\Phi} \\
	&= \sum_{i,j=1}^{n} \braket{\psi_i|\psi_j} \braket{\Phi|i} \braket{j|\Phi} \\
	&= \left\| \sum_{i=1}^{n} \braket{i|\Phi} \ket{\psi_i} \right\|^2 \\
	&\geq 0.
\end{align*}

The Gram matrix allows us to reformulate the exclusion problem in a more abstract, basis-independent manner. Since $\mathcal{G}$ is Hermitian and positive semidefinite, it can be factorized as,
\begin{align*}
	\mathcal{G} = X^\dagger X,
\end{align*}
for some matrix $X$, whose columns are the pure states,
\begin{align*}
	X = \begin{pmatrix}
		\mid & \mid &        & \mid \\
		\ket{\psi_1} & \ket{\psi_2} & \dots & \ket{\psi_n} \\
		\mid & \mid &        & \mid
	\end{pmatrix}.
\end{align*}

>>>>

\documentclass[12pt,letterpaper]{article}
\usepackage{fullpage}
\usepackage[top=1.5cm, bottom=3.5cm, left=2.2cm, right=2.2cm]{geometry}
\usepackage{amsmath,amsthm,amsfonts,amssymb,amscd, esint}
\usepackage{lastpage}
\usepackage{enumerate}
\usepackage{fancyhdr}
\usepackage{mathrsfs}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{hyperref}
\usepackage[english]{babel}
\usepackage{lipsum}
\usepackage[table,xcdraw]{xcolor}
\usepackage{enumitem}
\usepackage{overpic}
\usepackage{float}
\usepackage{yfonts}
\usepackage{braket}
\usepackage{dsfont}
\usepackage{tikz}
\usepackage{wrapfig}
\usepackage{url}
\usepackage{natbib}
\usepackage[normalem]{ulem}
\usepackage{multicol}
\useunder{\uline}{\ul}{}

@@

% --- Section formatting ---
\titleformat{\section}
  {\normalfont\large\bfseries\centering} % font size, bold, centered
  {\thesection}{1em}{\MakeTextUppercase} % uppercase text

@@

% --- Roman numerals for numbering ---
\renewcommand{\baselinestretch}{1.1} 
%\renewcommand{\thesection}{\Roman{section}}
%\renewcommand{\thesubsection}{\Roman{section}.\Roman{subsection}}

@@

\usepackage[skip=10pt , indent=20pt]{parskip}
%\setlength{\parindent}{0.0in}
%\setlength{\parskip}{0.05in}

% Edit these as appropriate
\newcommand\course{}
\newcommand\subject{End of Degree Project}
\newcommand\degree{Bachelor's Degree in Physics}
\newcommand\documenttitle{Lower bounds of the success probability in quantum state exclusion for general ensembles}
\newcommand\NetIDb{Universitat Autònoma de Barcelona}
\newcommand\AuthorName{Sergio Castañeiras Morales}

@@

\begin{document}
\title{\vspace{4cm} \normalsize 
		\includegraphics[width = 0.25\textwidth]{GeneralSources/UABLogo.png}\\ [0.5cm]
		\textsc{\NetIDb}\\ [2.0cm]
		\HRule{0.5pt} \\
		\LARGE \textbf{\uppercase{\documenttitle}}
		\HRule{2pt} \\ [1.5cm]
		\normalsize \begin{tabular}{rcl}  % Create a right-left column alignment
        \textsc{Author} & : & \textsc{\AuthorName} \\
        \textsc{Supervisor} & : & \textsc{Ramón Muñoz Tapia} \\
        \textsc{Co-Supervisor} & : & \textsc{Santiago Llorens Fernández}
    \end{tabular}
    \normalsize \vspace*{4.5\baselineskip}
		}

@@

\thispagestyle{empty}
\mbox{} 
\newpage
\thispagestyle{empty}
\vspace*{\fill} % Push content to vertical center
\begin{flushright}
    \emph{Ab ovo usque ad mala.}\\[1em]
    \textbf{Horace}
\end{flushright}
\vspace*{\fill} 
\newpage
\thispagestyle{empty}
\mbox{} 
\newpage
\thispagestyle{empty}
\cfoot{}
\tableofcontents

\newpage
\pagestyle{fancyplain}
\headheight 35pt
\lhead{\NetIDb}    
\rhead{\subject}
\cfoot{\small\thepage}
\headsep 1.5em
\setcounter{page}{1}
\begin{center}
\begin{abstract}
Given a quantum state known to be prepared from an ensemble of two or more states, quantum state exclusion aims to rule out the possibility that it was prepared in a particular state from the ensemble. Using the known solution for group generated ensembles \cite{MainPaper}, we study this result as a lower bound for randomly generated ensembles via semidefinite programming.
\end{abstract}
\end{center}
%Keywords: \emph{SDP, Quantum state exclusion , \textcolor{blue}{Add keywords}}.
\section{Introduction}

\hspace{20pt}In many real-world scenarios, excluding a certain hypothesis can be more practical than solving the problem entirely. For instance, in disease diagnosis, ruling out potential illnesses is often the first step toward identifying the actual condition. Similarly, in the context of detecting a gas leak in a complex pipe system, excluding the possibility of the gas being chlorine, a toxic substance, can guide the search for the faulty pipe by narrowing down the potential sources and providing critical information about the gas's toxicity. In numerous everyday situations, the strategy of excluding specific possibilities forms a natural and often effective initial approach to the problem.

In the quantum realm, a similar line of reasoning can be applied in some problmes. One of the fundamental tasks in quantum information theory is the identification of an unknown quantum state known to belong to a given ensemble of possible states. This task is referred to as Quantum State Discrimination (\gls{qsd}), and it has been extensively studied in the literature \cite{DiscriminationArticle}. However, instead of aiming to identify the exact state, we may take another approach by applying the hypothesis exclusion idea. That is, given a quantum state known to be one among a certain ensemble, the goal becomes to conclusively rule out the possibility of the unknown state being a specific member of that ensemble. This alternative task is known as Quantum State Exclusion (\gls{qse}) \cite{PhysicsExclusionSource}, and it offers a distinct framework with its own set of advantages \cite{piani2020quantum} with respect to its counterpart \gls{qsd}.

We can approach \gls{qse} task under several protocols, depending on the context within we perform the hypothesis exclusion we can prefer one or another. In this project, we focus on two main protocols, Minimum Error (\gls{me}) and Zero Error (\gls{ze})\footnote{Also known as \emph{unambiguous exclusion}.}, i.e. the protocols for which analytical results have been presented \cite{MainPaper}. In the Minimum Error scenario, the goal is to minimize the probability of mistakenly excluding the actual prepared state. In contrast, the Zero Error approach seeks to exclude a state with absolute certainty, although this procedure sometimes yields an inconclusive result when perfect exclusion cannot be accomplished.

While \gls{qsd} has seen significant progress over the years \cite{bae2015quantum, helstromBook}, \gls{qse} has remained relatively underexplored until recently. New analytical results have emerged \cite{MainPaper} for both the \gls{me} and \gls{ze} protocols, particularly when the ensemble of quantum states possesses a certain degree of symmetry. In particular, when the ensemble is generated through the action of a finite algebraic group, the problem becomes more tractable, highlighting \gls{qse} as a worthwhile object of study.

Although the tasks of exclusion and discrimination coincide for ensembles containing only two states\footnote{Since for the two states case excluding one necessarily implies identifying the other.}, when dealing with ensembles of three or more states, the two problems diverge in both approach and complexity. Though this work primarily focuses on \gls{qse}, we include a comparative discussion of both tasks and protocols to further motivate the interest in \gls{qse} and discuss the main differences with \gls{qsd} in Appendix \ref{appendixComparisonQSDvsQSE}. One of the key features of \gls{qse} is the possibility of achieving \emph{perfect exclusion}, i.e. confidently ruling out a candidate state with complete certainty that it is not the unknown state. In contrast, \emph{perfect discrimination} is generally not achievable, meaning that one cannot identify the unknown state with absolute certainty in most settings\cite{OptimalitySRM}.

This capacity for perfect exclusion opens new avenues in quantum information theory, especially in scenarios involving partial information retrieval from quantum systems \cite{gour2020weight}. By eliminating certain hypotheses, one can extract meaningful information from a quantum system without requiring full state identification.

Building upon the recent results for the probability of a correct exclusion for group-generated ensembles \cite{MainPaper},\footnote{A proper definition of a \emph{group generated ensemble} will be given in the \emph{Group generated ensembles} section \ref{sectionGroupGeneratedEnsemble}, for the time being, we can identify them as ensembles that comprehend set of states that exhibit a certain symmetry.} this project undertakes a numerical investigation into \gls{qse}, highlighting the analytical results from group-generated ensembles as lower bounds for more general cases. The project comences with a preliminaries section that outlines the mathematical framework underlying the \gls{qse} task, along with the essential tools used in our analysis. We then present results for the probability of correct hypothesis exclusion in ensembles consisting of three quantum states, the simplest case. Furthermore, we explore the distribution of correct exclusion probabilities for randomly generated ensembles sampled according to Haar measure.\footnote{The importance of the generic ensembles generation being performed according to Haar measure is a key for the numerical analysis and it is discussed in the \emph{Methodology} section (\ref{sectionMethodology})}

To carry out our numerical analysis, we employ \gls{sdp} techniques, which allow us to cast the \gls{qse} task as a convex optimization problem. The \gls{sdp} approach has proven particularly effective for problems of this kind \cite{bandyopadhyay2014conclusive}. Finally, ve conclude by providing some theoretical intuition and insights based on our results.

\newpage
\section{Preliminaries}

In this section we introduce some mathematical formalism of the postulates of Quantum Mechanics as well as some relevant objects in order to introduce the mathematical framework of \gls{qse} in a \gls{sdp} form. Additionally we define the Haar measure and discuss its importance for a proper generation of generic ensembles.

\subsection{Postulates and definitions}

This project is entirely situated within the quantum realm. Therefore, we begin by contextualizing the foundational postulates, notation, and elements upon which the rest of the project relies on.

@@

%Note that if $\ket{\psi'} = e^{i\theta}\ket{\psi}$ for some global phase $\theta \in \mathbb{R}$, the observable outcomes and probabilities remain unchanged. Therefore, quantum states differing only by a global phase are physically indistinguishable.

In relation to Postulate I, it is often convenient to represent quantum states using the \emph{density matrix} formalism, where a pure  state is described by $\rho = \ket{\psi}\bra{\psi}$ instead of $\ket{\psi}$.

%In this formalism, the global phase ambiguity disappears, since if $\ket{\psi'} = e^{i\theta}\ket{\psi}$, then
%\begin{align*}
%	\rho = \ket{\psi'}\bra{\psi'} = \ket{\psi}\bra{\psi} .
%\end{align*}

@@

Both formulations are equivalent, though one may be more convenient than the other depending on the context. Furthermore, the density matrix formalism naturally introduces the concept of \emph{mixed states}. Until now, we have implicitly dealt only with \emph{pure states}, i.e., states of the form $\rho = \ket{\psi}\bra{\psi}$. Suppose now that, for reasons we ignore, the system is prepared in one of several pure states $\{\rho_i\}_{i=1}^n$ with probabilities $\{\eta_i\}_{i=1}^n$. Then the overall state of the system is described by the mixed state
\begin{align*}
	\rho = \sum_{i=1}^n \eta_i \rho_i.
\end{align*}

@@

In relation to the observable postulate, we now define a \gls{povm} as a collection of measurement operators $\{\Pi_i\}_{i=1}^n$ that satisfy,
\begin{itemize}
	\item Each $\Pi_i$ is positive semi-definite hermitian operator, meaning all its eigenvalues $\pi_j^{(i)}$ satisfy $\pi_j^{(i)} \geq 0$.
	\item The operators satisfy the completeness relation,
\begin{align*}
	\sum_{i=1}^n \Pi_i = \mathds{1},
\end{align*}
where $\mathds{1}$ denotes the identity operator on the Hilbert space $\mathcal{H}$.
\end{itemize}

Unlike projective measurements, the operators $\Pi_i$ in a POVM are not necessarily orthogonal projectors. This allows the number of measurement outcomes to exceed the dimensionality of the Hilbert space, extending the previos orthogonal measurement to a more generic status and making POVMs particularly useful in tasks such as \gls{qse} and \gls{qsd}.

\subsection{Formulation of the problem}\label{sectionFormulationOfTheProblem}

Let $\left\{(\rho_i, \eta_i)\right\}_{i=1}^n$ be an ensemble of $n$ quantum states, where each $\rho_i$ denotes a pure state density matrix\footnote{This formulation hols true for mixed states but the project will only discuss the pure state scenario.}, i.e., $\rho_i = \ket{\psi_i}\bra{\psi_i}$, and $\eta_i$ represents the prior probability of occurrence of the state $\rho_i$. Let $\tilde{\rho}$ be a state from this ensemble. Our objective is to develop a procedure to identify another state $\rho_j \in \left\{\rho_i\right\}_{i=1}^n$, such that $\rho_j \neq \tilde{\rho}$.

Quantum measurements are described by a set of \glspl{povm}, denoted by $\left\{\Pi_i\right\}_{i=1}^n$, acting on the Hilbert space $\mathcal{H}$ of the quantum system. Here we present the two studied protocols for \gls{qse}: minimum-error (\gls{me}) and zero-error (\gls{ze}).

As previously discussed the goal of the \gls{me} protocol is to minimize the probability of incorrectly excluding the unknown state from our hypothesis. If we formulate it as an \gls{sdp}, the \gls{sdp} minimization of the error probability reads,\footnote{Note that the \gls{sdp} formulations of quantum state discrimination may differ from the exclusion ones by interchanging minimization and maximization problems.}
\begin{align*}
	P_{\text{\gls{me}}}^e = \min_{\left\{\Pi_i\right\}} &\sum_{i=1}^n \eta_i \tr(\Pi_i \rho_i),\\
	\text{subject to} \quad & \sum_{i=1}^n \Pi_i = \mathds{1}, \quad \Pi_i \geq 0 \quad \forall i \in \{1, \dots, n\}.
\end{align*}

Note the constraints $\sum_{i=1}^n \Pi_i = \mathds{1}$ and $\Pi_i \geq 0$ ensure that the set $\{\Pi_i\}_{i=1}^n$ form a valid \gls{povm}, raising a proper probability distribution. The superscript $e$ in $P_{\text{\gls{me}}}^e$ indicates that this is the \emph{error probability}.

Alternatively, we may formulate the problem in terms of the \emph{success probability}, denoted by $P_{\text{\gls{me}}}^s$, which quantifies the probability of a correct exclusion. This equivalent formulation reads,
\begin{align*}
	P_{\text{\gls{me}}}^s = \max_{\left\{\Pi_i\right\}} &\left(1 - \sum_{i=1}^n \eta_i\tr(\Pi_i \rho_i)\right),\\
	\text{subject to} \quad & \sum_{i=1}^n \Pi_i = \mathds{1}, \quad \Pi_i \geq 0 \quad \forall i \in \{1, \dots, n\}.
\end{align*}

Naturally, both formulations are related via,
\begin{align*}
P_{\text{\gls{me}}}^s + P_{\text{\gls{me}}}^e = 1.
\end{align*}

In the case of the \gls{ze} protocol, the objective is to exclude with total accuracy a state known to be different from the unkown state. However, the protocol might lead to inconclusive result. In order to do so, the \gls{povm}s must also satisfy an unambiguity condition, i.e. each measurement operator $\Pi_i$ must be orthogonal to the corresponding state $\rho_i$. In other words,
\begin{align*}
\tr(\Pi_i \rho_i) = 0 \quad \forall i \in \{1, \dots, n\}.
\end{align*}

To ensure completeness, we introduce an additional \gls{povm} element $\Pi_?$ representing an inconclusive result,
\begin{align*}
\Pi_? = \mathds{1} - \sum_{i=1}^n \Pi_i.
\end{align*}
If the measurement yields the outcome $\Pi_?$, i.e., the "?" label, the result is inconclusive.

The corresponding \gls{sdp} for minimizing the probability of an inconclusive result, i.e. error, in the \gls{ze} protocol is,
\begin{align*}
	P_{\text{\gls{ze}}}^e = \min_{\left\{\Pi_i\right\}} &\sum_{i=1}^n \eta_i\tr(\Pi_? \rho_i),\\
	\text{subject to} \quad & \sum_{i=1}^n \Pi_i + \Pi_? = \mathds{1}, \quad \Pi_? \geq 0,\\
	& \tr(\Pi_i \rho_i) = 0, \quad \Pi_i \geq 0 \quad \forall i.
\end{align*}

The corresponding success probability is naturally given by, the maximization of $1-P_{\text{\gls{ze}}}^e$ under the same constraints.

Notice this formulation is analogous to the \gls{me} protocol, with the crucial difference being the constraint $\tr(\Pi_i \rho_i) = 0$, enforcing unambiguous discrimination.

Additionally, in some cases it is useful to consider the dual formulation of the problem, as solving the dual version of the \gls{sdp} can often be more computationally efficient. Moreover, one may consider an ansatz, i.e. a candidate set of \gls{povm}s, and compute the exclusion success probability under a given protocol. This provides a lower bound for the success probability, since in principle we cannot guarantee that this ansatz is optimal, i.e. there may exist another set of \gls{povm}s yielding a higher success rate.

However, by evaluating the same set of \gls{povm}s within the dual formulation, we can obtain an upper bound on the success probability. If the upper and lower bounds coincide, we can conclude that the chosen \gls{povm} set is indeed optimal. For this reason, we present the dual formulation of the problem in Appendix~\ref{sectionDualFormulationOfTheProblem}. This procedure is standard for verifying the optimality of a given \gls{povm} set and closely resembles the internal computations performed by a \gls{sdp} solver.

\subsection{Gram Matrix Formulation}\label{sectionGramMatrixFormulation}
The Gram matrix formulation recasts the \gls{sdp} problems associated with the exclusion task in a more basis-independent manner. This approach reveals that the exclusion tasks for certain ensembles can be equivalent to those of others. As such, we adopt this formulation in our analysis, as it offers valuable insight into the underlying relationships between the quantum states in the ensemble.

Given an ensemble $\left\{(\rho_i, \eta_i)\right\}_{i=1}^n$ of $n$ quantum states we define the $\mathcal{G} \in \mathbb{C}^{n \times n}$ be the \emph{Gram matrix} of the ensemble, as the $n \times n$ positive semidefinite Hermitian matrix given by,
\begin{equation*}
	\mathcal{G} =
	\begin{pmatrix}
		\eta_1 & \sqrt{\eta_1\eta_2}\braket{\psi_1|\psi_2} & \dots & \sqrt{\eta_1\eta_n}\braket{\psi_1|\psi_n} \\
		\sqrt{\eta_2\eta_1}\braket{\psi_2|\psi_1} & \eta_2 & \dots & \sqrt{\eta_2\eta_n}\braket{\psi_2|\psi_n} \\
		\vdots & \vdots & \ddots & \vdots \\
		\sqrt{\eta_n\eta_1}\braket{\psi_n|\psi_1} & \sqrt{\eta_n\eta_2}\braket{\psi_n|\psi_2} & \dots & \eta_n
	\end{pmatrix},
\end{equation*}
that is, $\mathcal{G}_{i,j} = \sqrt{\eta_i\eta_j}\braket{\psi_i|\psi_j}$. Since all states are normalized, the diagonal elements are equal to $\eta_i$.

By construction, the Gram matrix is Hermitian,
\begin{align*}
	\mathcal{G}_{i,j}^* = \left( \sqrt{\eta_i\eta_j}\braket{\psi_i|\psi_j}\right)^* =  \sqrt{\eta_j\eta_i}\braket{\psi_j|\psi_i} = \mathcal{G}_{j,i}.
\end{align*}
Additionally, $\mathcal{G}$ is positive semidefinite. To see this, consider an arbitrary state $\ket{\Phi}$ and an orthonormal basis $\{\ket{i}\}_{i=1}^n$, then,
\begin{align*}
	\braket{\Phi|\mathcal{G}|\Phi} &= \braket{\Phi| \left( \sum_{i,j=1}^{n} \sqrt{\eta_i\eta_j}\braket{\psi_i|\psi_j} \ket{i}\bra{j} \right) |\Phi} \\
	&= \sum_{i,j=1}^{n} \sqrt{\eta_i\eta_j}\braket{\psi_i|\psi_j} \braket{\Phi|i} \braket{j|\Phi} \\
	&= \left\| \sum_{i=1}^{n} \eta_i\braket{i|\Phi} \ket{\psi_i} \right\|\\
	&\geq 0.
\end{align*}

Since $\mathcal{G}$ is Hermitian and positive semidefinite, it can be factorized as,
